---
title: "Agnia Vibriani"
output:
  word_document: default
  html_document: default
  pdf_document: default
date: "2023-10-02"
---
# Task 4
## Using R example datasets
**Describe briefly the content of the CO2 dataset using the help function**
```{r}
#load the data
data("CO2")
?CO2
```
The experiment is analyzing the cold tolerance of the grass species *Echinochola cruss-galli* by measuring the CO2 response curve for photosynthesis. The data contain the CO2 uptake at several levels of CO2 concentration in the chilled and non-chilled plants.

**What is the average and median CO2 uptake of the plants from Quebec and Missisipi?**
```{r}
library(tidyverse)
library(dplyr)
```

```{r}
data(CO2)
#select the plant originated from quebec only
Quebec <-filter(CO2, Type=="Quebec")
summary(Quebec)
#select the plant originated from Mississippi only
Mississippi<-filter(CO2, Type=="Mississippi")
summary(Mississippi)
```


```{r}
#generate the table
data <-data_frame(
  Plant_Origin = c("Quebec", "Mississippi"),
  Mean = c(33.54, 20.88),
  Median = c(37.15, 19.30))
knitr::kable(data, caption = "CO2 uptake based on plant origin")
```

**[OPTIONAL]In the airway example data from Bioconductor, how many genes are expressed in each sample? How many genes are ot expressed in any sample?** 
