---
title: "Task 4"
output:
  word_document: default
  html_document: default
  pdf_document: default
date: "2023-10-02"
---

# Using R example datasets

**1.Describe briefly the content of the CO2 dataset using the help function**

```{r}
#load the data
data("CO2")
?CO2
```

The experiment is analyzing the cold tolerance of the grass species *Echinochola cruss-galli* by measuring the CO2 response curve for photosynthesis. The data contain the CO2 uptake at several levels of CO2 concentration in the chilled and non-chilled plants.

**2.What is the average and median CO2 uptake of the plants from Quebec and Missisipi?**

```{r, results=FALSE}
library(tidyverse)
library(dplyr)
```

```{r}
data(CO2)
#select the plant originated from quebec only
Quebec <-filter(CO2, Type=="Quebec")
summary(Quebec)
#select the plant originated from Mississippi only
Mississippi<-filter(CO2, Type=="Mississippi")
summary(Mississippi)
```

```{r}

```
Another way is using tapply
```{r}
tapply(CO2$uptake, CO2$Type, mean)
```
```{r}
tapply(CO2$uptake, CO2$Type, median)
```

**[OPTIONAL]In the airway example data from Bioconductor, how many genes are expressed in each sample? How many genes are not expressed in any sample?**
