---
title: "Task-5"
output: pdf_document
date: "2023-10-03"
---

#R Functions

**1. Write a function that calculates the ratio of the mean and the median of a given vector**

*symmetrically distributed if the ratio of the mean and median close to 1*

```{r}
x <- c(1:10)
ratio_mean_median <- function(x){
  mean_value <- mean(x, na.rm=TRUE)
  median_value <- median(x, na.rm=TRUE)
  ratio <- mean_value/median_value
  ratio
}
ratio_mean_median(x)
#na.rm=TRUE to handle missing value
```

*right-skewed data if the ratio of the mean and median is greater than 1*

```{r}
y <- c(1,2,3,4,50)
ratio_mean_median(y)
```

*left-skewed data if the ratio of the mean and median is less than 1*

```{r}
z <-c(-2,30,31,32,33,34)
ratio_mean_median(z)
```

**2. Write a function that ignores the lowest and the highest value from a given vector and calculate the mean**

```{r}
x <-c(1:6)
mean_without_max_min <-function(x){
  ignore_max_min <- x[c(-which.max(x), -which.min(x))]
  result<-mean(ignore_max_min)
  result
}
mean_without_max_min(x)
```

**Short explanation about piping**

The pipe operator chains together a sequence of operation by passing the results as an input to the next operation, in a natural and logical order of thinking. The pipes are not recommended to use when the operation is simple (might add unnecessary complexity), too long (hard to debug), has multiple output/input, and the sequence doesn't match the logic.

**Short explanation about apply-family of functions (apply, lapply, sapply)**

These functions apply a specified function to selected rows or columns of matrices, data frame, and list. They offers flexibility and efficiency when dealing with large datasets, particularly during data preprocessing such as data cleaning, normalization, or summarizing the descriptive statistics.
